Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-10 paper/paper.tex
proof marked?
review done up to 4.1 end

\subsection{Correctness of liveness analysis}

@@

\section{Computing liveness and its encoding as DFA}\label{sec:computing}
Section~\ref{sec:liveness} gave  a context-sensitive liveness analysis
and  {\color  {red}  proved  it  correct  with  reference  to  a  {\em
    minefield} semantics}.  In  particular, $\mathcal{L}$ and \pmap{f}
together described a liveness set for each program point in a function
in terms of  a given $\sigma$ and \Lfonly.  However,  we still have to
describe how to obtain demand transformers \Lfonly\ from the rule {\sc
  live-define} and how to compute the specific demand $\sigma$ on each
function.

To  answer these  questions, it  is convenient  to cast  the equations
arising out of $\mathit{ref}$  and $\mathcal{L}$ as the rules of a
grammar.   But first  we modify  the rules  themselves to  a different
form.

@@

The  solution  is  to  also   treat  the  operations  mentioned  above
symbolically.  We  introduce three new symbols:  \bcar, \bcdr, \clazy.
These symbols are defined as a relation $\hookrightarrow$ between sets
of access paths.  First, \bcar\ and \bcdr\ are defined as:

\begin{align*}
  &\bcar\sigma \hookrightarrow \sigma' \mbox{ where } \sigma' = \{\alpha \mid \acar\alpha \in \sigma\}\\
  &\bcdr\sigma \hookrightarrow \sigma' \mbox{ where } \sigma' = \{\alpha \mid \acdr\alpha \in \sigma\}
\end{align*}
Thus \bcar\ selects those entries in $\sigma$ that have leading \acar, and removes the leading \acar\ from the.
The symbol \clazy\ reduces the set of strings following it to a set containing only $\epsilon$. It filters out, however, the empty set of strings.
\begin{align*}
  \clazy\sigma \hookrightarrow & \left\{ 
  \begin{array}{ll}
    \emptyset&\mbox{if}~\sigma = \emptyset\\
    \{\epsilon\} & \mbox{otherwise}
  \end{array}\right.
\end{align*}
We can  now rewrite the \CONS\  and the \CAR\  rules of $\mathit{ref}$
as:
\begin{align*}
&\mathit{ref\/}((\CONS~x~y),\sigma,\Lfonly)
= x.\bcar\sigma \cup y.\bcdr\sigma  \label{eqn:mod-cons},~
\mbox{and} \\
&\mathit{ref\/}((\CAR~x),\sigma,\Lfonly)
          =   x.\clazy\sigma \cup x.0\sigma
\end{align*}
and the \Lfunonly\ rule
for \SIF\ as:
\begin{align*}
\mathcal{L}((\SIF~x~e_1~e_2),\sigma,\Lfonly) =
                    &\mathcal{L}(e_1,\sigma,\Lfonly)~\uplus
        \mathcal{L}(e_2,\sigma,\Lfonly)~\uplus\\
        &[\,\mathit{\epath} \mapsto  x.\clazy \sigma]
\end{align*}
The rules for  \CDR, \PRIM\ and \NULLQ\ are also  modified in a manner
similar to \CAR.  %% The definitions  of \bcar\, \bcdr\ and \clazy\ induce
%% the following relation $\hookrightarrow$ over sets of access paths:
%% \begin{align*}
%%   &\sigma_1\bcar\sigma_2  \hookrightarrow
%%   \sigma_1\sigma_2'\mbox{ where } \sigma_2 = \acar\sigma_2',\\
%% &\sigma_1\bcdr\sigma_2  \hookrightarrow
%%   \sigma_1\sigma_2'\mbox{ where } \sigma_2 =
%%   \acdr\sigma_2',~\mbox{and}\\
%% &\sigma_1\clazy \sigma_2 \hookrightarrow \sigma_1, \mbox{ if }
%% \sigma_2 \ne
%%   \emptyset
%% \end{align*}
When  there  are  multiple  occurences of  \bcar,  \bcdr  and  \clazy,
$\hookrightarrow$  is  applied  from  right to  left.   The  reflexive
transitive   closure  of   $\hookrightarrow$   will   be  denoted   as
$\stackrel{*}{\hookrightarrow}$.   The  following proposition  relates
the original and the modified liveness rules.

\begin{proposition}
Assume that a liveness computation  based on the original set of rules
gives the liveness  of the variable $x$ at a  program point $\pi_i$ as
$\sigma$   (symbolically,   $\Lanv{i}{x}=   \sigma$).   Further,   let
$\Lanv{i}{x}=  \sigma'$ when the  modified rules  are used  instead of
\Lfunonly.  Then $\sigma' \stackrel{*}{\hookrightarrow} \sigma$.
\end{proposition}

@@

%Thick red line
%{\color{red} \medskip \hrule \hrule \hrule  \medskip }

\newcommand{\emm}[2]{\ensuremath{\mathcal{#1}_{#2}}}

\subsection{Generating Liveness Equations}

>>>>

{\color{red}
\subsection{Correctness of liveness analysis}

@@

\section{Computing liveness and its encoding as DFA}\label{sec:computing}
Section~\ref{sec:liveness} gave  a context-sensitive liveness analysis
and  {\color  {red}  proved  it  correct  with  reference  to  a  {\em
    minefield} semantics}.  In  particular, $\mathcal{L}$ and \pmap{f}
together described a liveness set for each program point in a function
in terms of  a given $\sigma$ and \Lfonly.  However,  we still have to
describe how to obtain demand transformers \Lfonly\ from the rule {\sc
  live-define} and how to compute the specific demand $\sigma$ on each
function.
%%
To  answer these  questions, it  is convenient  to cast  the equations
arising  out of  $\mathit{ref}$ and  $\mathcal{L}$ as  the rules  of a
grammar.   To do  so, we  need  to modify  the rules  themselves to  a
different form.

@@

The  solution  is  to  also   treat  the  operations  mentioned  above
symbolically.  We  introduce three new symbols:  \bcar, \bcdr, \clazy.
These symbols are defined as a relation $\hookrightarrow$ between sets
of access paths:
\begin{align*}
  &\bcar\sigma \hookrightarrow \sigma' \mbox{ where } \sigma' = \{\alpha \mid \acar\alpha \in \sigma\}\\
  &\bcdr\sigma \hookrightarrow \sigma' \mbox{ where } \sigma' = \{\alpha \mid \acdr\alpha \in \sigma\}
\end{align*}
Thus \bcar\ selects those entries in $\sigma$ that have leading \acar, and removes the leading \acar\ from them.
The symbol \clazy\ reduces the set of strings following it to a set containing only $\epsilon$. It filters out, however, the empty set of strings.
\begin{align*}
  \clazy\sigma \hookrightarrow & \left\{ 
  \begin{array}{ll}
    \emptyset&\mbox{if}~\sigma = \emptyset\\
    \{\epsilon\} & \mbox{otherwise}
  \end{array}\right.
\end{align*}
We can  now rewrite the \CONS\  and the \CAR\  rules of $\mathit{ref}$
as:
\begin{align*}
&\mathit{ref\/}((\CONS~x~y),\sigma,\Lfonly)
= x.\bcar\sigma \cup y.\bcdr\sigma  \label{eqn:mod-cons},~
\mbox{and} \\
&\mathit{ref\/}((\CAR~x),\sigma,\Lfonly)
          =   x.\clazy\sigma \cup x.0\sigma
\end{align*}
and the \Lfunonly\ rule
for \SIF\ as:
\begin{align*}
\mathcal{L}((\SIF~x~e_1~e_2),\sigma,\Lfonly) =
                    &\mathcal{L}(e_1,\sigma,\Lfonly)~\uplus
        \mathcal{L}(e_2,\sigma,\Lfonly)~\uplus\\
        &[\,\mathit{\epath} \mapsto  x.\clazy \sigma]
\end{align*}
The rules for  \CDR, \PRIM\ and \NULLQ\ are also  modified in a manner
similar to \CAR.

When  there  are  multiple  occurences of  \bcar,  \bcdr\  and  \clazy,
$\hookrightarrow$  is  applied  from  right to  left.   The  reflexive
transitive   closure  of   $\hookrightarrow$   will   be  denoted   as
$\stackrel{*}{\hookrightarrow}$.   The  following proposition  relates
the original and the modified liveness rules.
\begin{proposition}
Assume that a liveness computation  based on the original set of rules
gives the liveness  of the variable $x$ at a  program point $\pi_i$ as
$\sigma$   (symbolically,   $\Lanv{i}{x}=   \sigma$).   Further,   let
$\Lanv{i}{x}=  \sigma'$ when the  modified rules  are used  instead of
\Lfunonly.  Then $\sigma' \stackrel{*}{\hookrightarrow} \sigma$.
\end{proposition}

@@

\newcommand{\emm}[2]{\ensuremath{\mathcal{#1}_{#2}}}
\subsection{Generating Liveness Equations}
